---
title: "Linear Models and Regression Analysis Homework 4"
author: "Yifei Dong"
date: "9/23/2020"
output: pdf_document
---

# 0. Data Preparation
setwd("/Users/yifei/Documents/Teachers College/Linear Models and Regression/Week 4/hw4")
getwd()

# 1. Flavor Deterioration

The results shown below were obtained in a small-scale experiment to study the relation between $^{\circ}F$ of storage temperature ($X$) and number of weeks before flavor deterioration of a food product begins to occur ($Y$).

Use the data from Exercises 5.4 on p.210. Using \textbf{\emph{matrix methods}}, find the following:

```{r}
# Get data from Exercise 5.4 on p.210
mydata<-read.table(
  "http://users.stat.ufl.edu/~rrandles/sta4210/Rclassnotes/data/textdatasets/KutnerData/Chapter%20%205%20Data%20Sets/CH05PR04.txt")
# Renames the variable
names(mydata)<-c("Y","X")
mydata
```

$~$

(1). $Y'Y$

```{r}
# Create the design matrix Y
Y<-matrix(mydata$Y,ncol = 1,nrow(mydata))
Y
```

```{r}
# Create the design matrix X
X<-cbind(rep(1,nrow(mydata)),mydata$X)
X
```

```{r}
# Y'Y
t(Y) %*% Y
```

$~$

(2). $X'X$

```{r}
# X'X
t(X) %*% X
```

$~$

(3). $(X'X)^{-1}$

```{r}
#(X'X)^(-1)
solve(t(X) %*% X)
```

$~$

(4). Vector $\textbf{b}$ of estimated regression coefficients

```{r}
# b=((X'X)^(-1))X'Y
b<-solve(t(X) %*% X) %*% t(X) %*% Y
b
```

```{r}
# Check and compare to built-in lm function
reg<-lm(mydata$Y~mydata$X)
summary(reg)
```

$~$

(5). Vector of residuals

```{r}
# Predicted Values: Xb
y.hat<-X %*% b
y.hat
```

```{r}
# Cross-check with build-in function
cbind(predict(reg),y.hat)
```

The predicted values calculated using the matrix methods are equal to the the results with build-in function

```{r}
# Vector of Residuals
e<-Y-y.hat
e
# Cross-check with build-in function
residuals(reg)
```

$~$

```{r}
# Sample size
n<-nrow(mydata)
# Identity matrix
I<-diag(rep(1,n))
# H matrix: X(X'X)^(-1)X'
H<-X %*% solve(t(X) %*% X) %*% t(X)
# Matrix methods for Residuals
(I-H) %*% Y
```

$~$

(6). SSR

```{r}
# J matrix
J<-matrix(rep(1,n^2),n,n)
# SSR=Y'(H-(1/n)J)Y
SSR<-t(Y) %*% (H-(J/n)) %*% Y
SSR
```

$~$

(7). SSE

```{r}
# SSE=e'e=Y'Y-b'X'Y
SSE<-t(Y) %*% Y-t(b) %*% t(X) %*% Y
SSE
```

```{r}
# Check with ANOVA Table
anova(reg)
# The results are the same.
```

$~$

(8). Cov($b$)

```{r}
# Covariance matrix of b
# MSE
MSE<-SSE/(n-2)
MSE
```

```{r}
# cov(b)=MSE(X'X)^(-1)
s.sq.b<-drop(MSE)*solve(t(X) %*% X)
s.sq.b
```

$~$

(9). Point estimate of $E\{Y_{h}\}$ when $X_{h}=-6$

```{r}
x.new<-c(1,-6)
y.hat.h<-t(x.new) %*% b
y.hat.h
```

$~$

(10). Hat matrix $H$

```{r}
# H matrix: X(X'X)^(-1)X'
H<-X %*% solve(t(X) %*% X) %*% t(X)
H
```


