---
title: "HW5"
author: "Chengxuan Hu"
date: "10/7/2020"
output: word_document
---

p248 Q6.5
```{r}
 # Read Table6.5
 t6.5<-read.table("http://users.stat.ufl.edu/~rrandles/sta4210/Rclassnotes/data/textdatasets/KutnerData/Chapter%20%206%20Data%20Sets/CH06PR05.txt", header=FALSE)
 # Column names can be changed with
 names(t6.5)[1]<-paste("Y")
 names(t6.5)[2]<-paste("X1")
 names(t6.5)[3]<-paste("X2")
 
 # To use the column names without reference to t6.5 you need to attach the dataset:
 attach(t6.5)
```
a. Obtain the scatter plot matrix and the correlation matrix. What information do these diag- nostic aids provide here?
```{r}
# scatter plot
pairs(t6.5)

# correlation matrix
cor(t6.5)

## Y and X1 are highly positively correlated while Y and X2 are moderately positively correlated. There is no linear correlation between X1 and X2.
```
b. Fit regression model (6.1) to the data. State the estimated regression function. How is bi interpreted here?
```{r}
reg<-lm(Y ~ X1 + X2)
summary(reg)

# the estimated regression function:
## Y.hat = 37.65+4.425*X1+4.3750*X2

## b1 is the slope of X1, which means holding X2 constant, Y will increase by 4.425 with the increase one unit of X1.
## That is to say, holding sweetness constant, degree of brand liking will increase by 4.425 with the increase one unit of moisture content.
```

c. Obtain the residuals and prepare a box plot of the residuals. What information does this plot provide?
```{r}
e<-resid(reg)
boxplot(e)

## The residuals are symmetrically distributed around zero. There are no outliers.
```
d. Plot the residuals against Y, X1, X2, and X1X2 on separate graphs. Also prepare a normal probability plot. Interpret the plots and summarize your findings.
```{r}
Yhat<-predict(reg)
plot(Yhat,e)
plot(X1,e)
plot(X2,e)
plot(X1*X2,e)

## There are no outliers. Linearity is reasonable. Constant variance holds. No interaction effect.

qqnorm(e)
qqline(e)
## Normality holds.
```
e. Conduct the Breusch-Palgan test for constancy of the error variance. State the alternatives, decision rule, and conclusion.
```{r}
library(lmtest)
bptest(reg)

## H0: gamma1 = gamma2 = 0
## Ha: gamma1 or gamma2 is not zero, that is constant variance violates.
## decision rule: if p-value<0.01, reject H0
## Conlusion: p-value=0.3599>0.01, do not reject H0. That is, constant variance holds.
```

p245 Q6.27
```{r}
 # Read Table6.27
 t6.27<-read.table("http://users.stat.ufl.edu/~rrandles/sta4210/Rclassnotes/data/textdatasets/KutnerData/Chapter%20%206%20Data%20Sets/CH06PR27.txt", header=FALSE)
 # Column names can be changed with
 names(t6.27)[1]<-paste("y")
 names(t6.27)[2]<-paste("x1")
 names(t6.27)[3]<-paste("x2")
 
 attach(t6.27)
 
 Y = y
 X = cbind(rep(1, nrow(t6.27)), x1, x2)
```

(a)
```{r}
 (b = solve(t(X)%*%X)%*%t(X)%*%y)
```

(b)
```{r}
 y.hat = X%*%b

 (e = y - y.hat)
```

(c)
```{r}
(H = X %*% solve(t(X)%*%X) %*% t(X))
```

(d)
```{r}
 # J matrix
 n = nrow(t6.27)
 J = matrix(rep(1, n^2), n, n)
 # SSR
 (SSR = t(y) %*% (H - J/n) %*% y)
```

(e)
```{r}
 SSE = t(y)%*%y - t(b)%*%t(X)%*%y
 MSE = SSE/(nrow(t6.27)-ncol(t6.27))

 (s.sq.b = drop(MSE)*solve(t(X)%*%X))
```

(f)
```{r}
Xh<-rbind(1,10,30)
t(Xh) %*% b
```

(g)
```{r}
drop(MSE)*(t(Xh)%*%solve(t(X) %*% X)%*%Xh)
```

