---
title: "HUDM5126 Linear Models and Regression Analysis Final Project"
author: "Yifei Dong"
date: "11/27/2020"
output: pdf_document
---


# 0. Data Preparation

```{r}
# Load packages
library(dplyr)
library(ggplot2)
library(GGally)
library(gridExtra)
library(qwraps2)
library(stargazer)
library(MASS)
library(bestglm)
library(car)
library(lmtest)
library(nortest)
```

```{r}
# Load Data
load("movies.Rdata")
# First 5 observations of the dataset
head(movies, 5)
```

```{r}
# Generalizability
# summary counts of movies by year
mydata <- movies %>%
  group_by(thtr_rel_year) %>%
  tally()
mydata
```

```{r}
# Number of movies released in theater by year
g1 <- ggplot(movies, aes(x = thtr_rel_year, y = ..count..))+
  geom_bar(fill = "gray25", color = "white")+
  scale_x_continuous("Year the movie is released in theaters")+
  scale_y_continuous("Count")+
  ggtitle("Number of Movies released in theater per year")+
  theme(plot.title = element_text(size = 12, face = "bold", hjust = 0.5))
g1
```

```{r}
# Box plot of number of movies released in theater per year 
# Distribution of the number of movies released in theater per year
g2 <- ggplot(mydata, aes(y = n))+
  stat_boxplot(geom = "errorbar", width = 0.15)+
  geom_boxplot()+
  theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())+
  scale_y_continuous("Count")
g2
```

Based on the summary statistics and box plot of the count, 75$\%$ of the years in the data set have less than 20 movies
released. Based on the rule of thumb, we need at least 30 observations to generalize our results to the whole 
population. Therefore, we could not generalize our research results to all U.S movies produced and released in each 
year. However, due to the fact that this data set consists of 651 randomly sampled movies, and the total number of 
observations (651) in this data set is well below 10$\%$ of all movies produced and released in the United Sates before
2016, we could safely generalize this data to the population of interest.

```{r}
# Remove variables for informational purposes and time relevant variables
movies2 <- movies %>%
  dplyr::select(title, title_type, genre, runtime, mpaa_rating, imdb_rating, imdb_num_votes,
         critics_rating, critics_score, audience_rating, audience_score, best_pic_nom, best_pic_win,
         best_actor_win, best_actress_win, best_dir_win, top200_box)
head(movies2, 5)
```

```{r}
# Check if the observations have NA values
movies2[!complete.cases(movies2), ]
```

There is one observation with NA value for `runtime`, which is the "The End of America". Remove this observations from the 
data set.

```{r}
# Remove rows with NA values
movies2 <- na.omit(movies2)
nrow(movies2) # We then have 650 observations
```

# 1. Exploratory Data Analysis

```{r}
# Summary statistics for predictor variables
summary(movies2[, -1])
```


```{r, results='hide'}
# LaTeX Descriptive Statistics Table for categorical variables
our_summary1 <- 
  list("Title Type" = 
         list("Documentary" = ~qwraps2::n_perc(title_type == "Documentary", 
                                               digits = 2),
              "Feature Film" = ~qwraps2::n_perc(title_type == "Feature Film", 
                                                digits = 2),
              "TV Movie" = ~qwraps2::n_perc(title_type == "TV Movie", 
                                            digits = 2)),
       "Genre" = 
         list("Drama" = ~qwraps2::n_perc(genre == "Drama", 
                                         digits = 2),
              "Comedy" = ~qwraps2::n_perc(genre == "Comedy", 
                                          digits = 2),
              "Action and Adventure" = ~qwraps2::n_perc(genre == "Action & Adventure", 
                                                      digits = 2),
              "Animation" = ~qwraps2::n_perc(genre == "Animation", 
                                             digits = 2),
              "Mystery and Suspense" = ~qwraps2::n_perc(genre == "Mystery & Suspense", 
                                                      digits = 2),
              "Documentary" = ~qwraps2::n_perc(genre == "Documentary", 
                                               digits = 2),
              "Horror" = ~qwraps2::n_perc(genre == "Horror", digits = 2),
              "Science Fiction and Fantasy" = ~qwraps2::n_perc(genre == "Science Fiction & Fantasy", 
                                                            digits = 2),
              "Musical and Performing Arts" = ~qwraps2::n_perc(genre == "Musical & Performing Arts", 
                                                            digits = 2),
              "Art House and International" = ~qwraps2::n_perc(genre == "Art House & International", 
                                                            digits = 2),
              "Other" = ~qwraps2::n_perc(genre == "Other", 
                                         digits = 2)),
       "MPAA Rating" = 
         list("G" = ~qwraps2::n_perc(mpaa_rating =="G", 
                                     digits = 2),
              "NC-17" = ~qwraps2::n_perc(mpaa_rating == "NC-17", 
                                         digits = 2),
              "PG" = ~qwraps2::n_perc(mpaa_rating == "PG", 
                                      digits = 2),
              "PG-13" = ~qwraps2::n_perc(mpaa_rating == "PG-13", 
                                         digits = 2),
              "R" = ~qwraps2::n_perc(mpaa_rating == "R", 
                                     digits = 2),
              "Unrated" = ~qwraps2::n_perc(mpaa_rating == "Unrated",
                                           digits = 2)),
       "Critics Rating" =
         list("Certified Fresh" = ~qwraps2::n_perc(critics_rating == "Certified Fresh", 
                                                   digits = 2),
              "Fresh" = ~qwraps2::n_perc(critics_rating == "Fresh", 
                                         digits = 2),
              "Rotten" = ~qwraps2::n_perc(critics_rating == "Rotten", 
                                          digits = 2)),
       "Audience Rating" = 
         list("Spilled" = ~qwraps2::n_perc(audience_rating == "Spilled",
                                           digits = 2),
              "Upright" = ~qwraps2::n_perc(audience_rating == "Upright", digits = 2)),
       "Best Picture Nomination" = 
         list("No" = ~qwraps2::n_perc(best_pic_nom == "no", 
                                      digits = 2),
              "Yes" = ~qwraps2::n_perc(best_pic_nom == "yes", 
                                       digits = 2)),
       "Won Oscar Best Picture Award" = 
         list("No" = ~qwraps2::n_perc(best_pic_win == "no", 
                                      digits = 2),
              "Yes" = ~qwraps2::n_perc(best_pic_win == "yes", 
                                       digits = 2)),
       "Main Actors ever won an Oscar" = 
         list("No" = ~qwraps2::n_perc(best_actor_win == "no", 
                                      digits = 2),
              "Yes" = ~qwraps2::n_perc(best_actor_win == "yes", 
                                       digits = 2)),
       "Main Actresses ever won an Oscar" = 
         list("No" = ~qwraps2::n_perc(best_actress_win == "no", 
                                      digits = 2),
              "Yes" = ~qwraps2::n_perc(best_actress_win == "yes", 
                                       digits = 2)),
       "Director ever won an Oscar" =
         list("No" = ~qwraps2::n_perc(best_dir_win == "no", 
                                      digits = 2),
              "Yes" = ~qwraps2::n_perc(best_dir_win == "yes", 
                                       digits = 2)),
       "Movie ever in the Top 200 Box Office" =
         list("No" = ~qwraps2::n_perc(top200_box == "no", 
                                      digits = 2),
              "Yes" = ~qwraps2::n_perc(top200_box == "yes", 
                                       digits = 2)))
categorical_summary <- summary_table(movies2, our_summary1)
```


## 1.1 Categoriacal Variables

```{r}
# title type
summary1 <- movies2 %>%
  group_by(title_type) %>%
  tally()
summary1
```

```{r}
# Bar plot of Type of Movie
g3 <- ggplot(summary1, aes(x = title_type, y = n))+
  geom_bar(stat = "identity", fill = "grey", color = "black")+
  scale_x_discrete("Type of Movie")+
  scale_y_continuous("Count")+
  geom_text(aes(label = n), vjust = -0.5, color = "black", size = 3.5, 
            face = "bold", fontface = "bold")+
  theme(axis.title = element_text(face = "bold", hjust = 0.5, size = 8))
g3
```

```{r}
# Genre
summary2 <- movies %>%
  group_by(genre) %>%
  tally()
summary2
```

```{r}
# Bar plot of Genre
g4 <- ggplot(summary2, aes(x = genre, y = n))+
  geom_bar(stat = "identity", fill = "grey", color = "black")+
  scale_x_discrete("Genre of Movie")+
  scale_y_continuous("Count")+
  coord_flip()+
  geom_text(aes(label = n), position = position_stack(vjust = 0.5), size = 3.5, fontface = "bold")+
  theme(axis.title = element_text(face = "bold", hjust = 0.5, size = 10))
g4
```

```{r}
# Critics Rating
summary3 <- movies2 %>%
  group_by(critics_rating) %>%
  tally()
summary3
```

```{r}
# Critics Rating
g5 <- ggplot(summary3, aes(x = critics_rating, y = n))+
  geom_bar(stat = "identity", fill = "grey", color = "black")+
  scale_x_discrete("Critics rating on Rotten Tomatoes")+
  scale_y_continuous("Count")+
  geom_text(aes(label = n), position = position_stack(vjust = 0.5), size = 3.5, fontface = "bold")+
  theme(axis.title = element_text(face = "bold", hjust = 0.5, size = 8))
g5
```

```{r}
# Best Picture Oscar nomination
g6 <- ggplot(movies2, aes(x = best_pic_nom, y = ..prop.., group = 1))+
  geom_bar(fill = "grey", color = "black")+
  scale_x_discrete("Oscar Best Picture Nomination", label = c("No", "Yes"))+
  scale_y_continuous("Percent", labels = scales::percent)+
  geom_text(aes(label = scales::percent(..prop.., accuracy = 0.01), y = (..prop..)), stat = "count", 
            position = position_stack(vjust = 0.5), size =3)+
  theme(axis.title = element_text(size = 8, hjust = 0.5, face = "bold"))
g6
```

```{r}
# Best Picture Won Status
g7 <- ggplot(movies2, aes(x = best_pic_win, y = ..prop.., group = 1))+
  geom_bar(fill = "grey", color = "black")+
  scale_x_discrete("Movie ever won an Oscar Best Picture", label = c("No", "Yes"))+
  scale_y_continuous("Percent", labels = scales::percent)+
  geom_text(aes(label = scales::percent(..prop.., accuracy = 0.01), y = (..prop..)), stat = "count", 
            position = position_stack(vjust = 0.5), angle = 35, size = 3)+
  theme(axis.title = element_text(size = 8, hjust = 0.5, face = "bold"))
g7
```

```{r}
# Actors in the movie ever won an Oscar
g8 <- ggplot(movies2, aes(x = best_actor_win, y = ..prop.., group = 1))+
  geom_bar(fill = "grey", color = "black")+
  scale_x_discrete("Actors ever won an Oscar", label = c("No", "Yes"))+
  scale_y_continuous("Percent", labels = scales::percent)+
  geom_text(aes(label = scales::percent(..prop.., accuracy = 0.01), y = (..prop..)), stat = "count", 
            position = position_stack(vjust = 0.5), size = 3)+
  theme(axis.title = element_text(size = 8, hjust = 0.5, face = "bold"))
g8
```

```{r}
# Actresses in the movie ever won an Oscar
g9 <- ggplot(movies2, aes(x = best_actress_win, y = ..prop.., group = 1))+
  geom_bar(fill = "grey", color = "black")+
  scale_x_discrete("Actresses ever won an Oscar", label = c("No", "Yes"))+
  scale_y_continuous("Percent", labels = scales::percent)+
  geom_text(aes(label = scales::percent(..prop.., accuracy = 0.01), y = (..prop..)), stat = "count", 
            position = position_stack(vjust = 0.5), size = 3)+
  theme(axis.title = element_text(size = 8, hjust = 0.5, face = "bold"))
g9
```

```{r}
# Director of the movie ever won an Oscar
g10 <- ggplot(movies2, aes(x = best_dir_win, y = ..prop.., group = 1))+
  geom_bar(fill = "grey", color = "black")+
  scale_x_discrete("Director ever won an Oscar", label = c("No", "Yes"))+
  scale_y_continuous("Percent", labels = scales::percent)+
  geom_text(aes(label = scales::percent(..prop.., accuracy = 0.01), y = (..prop..)), stat = "count", 
            position = position_stack(vjust = 0.5), size = 3)+
  theme(axis.title = element_text(size = 8, hjust = 0.5, face = "bold"))
g10
```

```{r}
# Movie ever in the Top 200 Box Office
g11 <- ggplot(movies2, aes(x = top200_box, y = ..prop.., group = 1))+
  geom_bar(fill = "grey", color = "black")+
  scale_x_discrete("Movie ever in the Top 200 Box Office", label = c("No", "Yes"))+
  scale_y_continuous("Percent", labels = scales::percent)+
  geom_text(aes(label = scales::percent(..prop.., accuracy = 0.01), y = (..prop..)), stat = "count", 
            position = position_stack(vjust = 0.5), size = 3)+
  theme(axis.title = element_text(size = 8, hjust = 0.5, face = "bold"))
g11
```

```{r}
grid.arrange(g6, g7, g8, g9, g10, g11, ncol = 3, nrow = 2)
```

```{r}
# MPAA rating of the Movie
g12 <- ggplot(movies2, aes(x = mpaa_rating))+
  geom_bar(aes(y = ..count..), fill = "grey", color = "black")+
  geom_text(aes(label = (..count..), y = (..count..)), stat = "count", 
            vjust = -0.5, size = 3, fontface = "bold")+
  scale_x_discrete("MPAA rating of the Movie")+
  scale_y_continuous("Count")+
  theme(axis.title = element_text(face = "bold", hjust = 0.5, size = 8))
g12
```

```{r}
# Audience Rating
g13 <- ggplot(movies2, aes(x = audience_rating))+
  geom_bar(aes(y = ..count..), fill = "grey", color = "black")+
  geom_text(aes(label = (..count..), y = (..count..)), stat = "count", 
            position = position_stack(vjust = 0.5), size = 3, fontface = "bold")+
  scale_x_discrete("Audience rating of the Movie")+
  scale_y_continuous("Count")+
  theme(axis.title = element_text(face = "bold", hjust = 0.5, size = 8))
g13
```

```{r}
grid.arrange(g3, g5, g12, g13, ncol = 2)
```

## 1.2 Numerical Variables

### 1.2.1 Response Variable

```{r}
# Histogram of response variable: Audience Score
g14 <- ggplot(movies2, aes(x = audience_score))+
  geom_histogram(fill = "white", col = "black")+
  scale_x_continuous("Audience Score")+
  scale_y_continuous("Frequency")+
  geom_vline(xintercept = mean(movies2$audience_score), linetype = "dashed", col = "red")+
  geom_vline(xintercept = median(movies2$audience_score), linetype = "dashed", col = "blue")+
  geom_text(label = "Mean: 62.35", x = 55, y = 45, col = "red")+
  geom_text(label = "Median: 65", x = 72, y = 45, col = "blue")
g14
```

```{r}
summary(movies2$audience_score)
```

The median of the response variable, audience score is 65. 25% of the movies in the data set have 
audience score higher than 80 and 25% of the movies have audience score lower than 46. The lowest 
audience score in the data set is 11 and the highest is 97. Also, due to the fact that mean of audience 
score is less than the median of audience score, we have a left-skewed distribution.

```{r}
# Movie Run time
g15 <- ggplot(movies2, aes(x = runtime))+
  geom_histogram(fill = "white", color = "black")+
  scale_y_continuous("Count")+
  scale_x_continuous("Run time")
g15
```

```{r}
# IMDB rating
g16 <- ggplot(movies, aes(x = imdb_rating))+
  geom_histogram(fill = "white", color = "black")+
  scale_x_continuous("IMDB rating")+
  scale_y_continuous("Count")
g16
```

```{r}
g17 <- ggplot(movies2, aes(x = imdb_num_votes))+
  geom_histogram(fill = "white", color = "black")+
  scale_x_continuous("IMDB Number of Votes")+
  scale_y_continuous("Count")
g17
```

```{r}
# Critics scores
g18 <- ggplot(movies2, aes(x = critics_score))+
  geom_histogram(fill = "white", color = "black")+
  scale_x_continuous("Critics Scores")+
  scale_y_continuous("Count")
g18
```

```{r}
grid.arrange(g15, g16, g17, g18, ncol = 2)
```

```{r}
summary(movies2$runtime) # Run time
summary(movies2$imdb_rating) # IMDB rating
summary(movies2$imdb_num_votes) # IMDB Number of Votes
summary(movies2$critics_score) # Critics Scores on Rotten Tomatoes
```

The run time, IMDB number of votes have a right-skewed distribution, while IMDB rating and critics scores on Rotten Tomatoes have left-skewed distributions.

### 1.2.2 Response Variable and Categorical Variables

```{r}
# Distribution of Audience Scores by Movies Genre
g19 <- ggplot(movies2, aes(x = genre, y = audience_score))+
  stat_boxplot(geom = "errorbar")+
  geom_boxplot()+
  scale_x_discrete("Genre", 
                   labels = c("Action & Adventure" = "A & A", "Art House & International" = "AH & I", 
                              "Musical & Performing Arts" = "M & PA", "Mystery & Suspense" = "M & S",
                              "Science Fiction & Fantasy" = "SF & F"))+
  scale_y_continuous("Audience Score")
g19
```

```{r}
# Distribution of Audience Score by MPAA Rating
g20 <- ggplot(movies2, aes(x = mpaa_rating, y = audience_score))+
  stat_boxplot(geom = "errorbar", width = 0.15)+
  geom_boxplot()+
  scale_x_discrete("MPAA Rating")+
  scale_y_continuous("Audience Score")
g20
```

```{r}
# Distribution of Audience Score by Oscar Best Picture Nomination Status
g21 <- ggplot(movies2, aes(x = best_pic_nom, y = audience_score))+
  stat_boxplot(geom = "errorbar", width = 0.15)+
  geom_boxplot()+
  scale_x_discrete("Oscar Best Picture Nomination", label = c("No", "Yes"))+
  scale_y_continuous("Audience Score")
g21
```

```{r}
# Distribution of Audience Score by Oscar Best Picture Award Status
g22 <- ggplot(movies2, aes(x = best_pic_win, y = audience_score))+
  stat_boxplot(geom = "errorbar", width = 0.15)+
  geom_boxplot()+
  scale_x_discrete("Movie ever won an Oscar Best Picture", labels = c("No", "Yes"))+
  scale_y_continuous("Audience Score")
g22
```

```{r}
# Distribution of Audience Score by Actors ever won an Oscar
g23 <- ggplot(movies2, aes(x = best_actor_win, y = audience_score))+
  stat_boxplot(geom = "errorbar", width = 0.15)+
  geom_boxplot()+
  scale_x_discrete("Actors ever won an Oscar", labels = c("No", "Yes"))+
  scale_y_continuous("Audience Score")
g23
```

```{r}
# Distribution of Audience Score by Actresses ever won an Oscar
g24 <- ggplot(movies2, aes(x = best_actress_win, y = audience_score))+
  stat_boxplot(geom = "errorbar", width = 0.15)+
  geom_boxplot()+
  scale_x_discrete("Actresses ever won an Oscar", labels = c("No", "Yes"))+
  scale_y_continuous("Audience Score")
g24
```

```{r}
# Distribution of Audience Score by Director ever won an Oscar
g25 <- ggplot(movies2, aes(x = best_dir_win, y = audience_score))+
  stat_boxplot(geom = "errorbar", width = 0.15)+
  geom_boxplot()+
  scale_x_discrete("Movie Director has ever won an Oscar", labels = c("No", "Yes"))+
  scale_y_continuous("Audience Score")
g25
```

```{r}
# Distribution of Audience Score by Movie has ever in the Top 200 Box Office
g26 <- ggplot(movies2, aes(x = top200_box, y = audience_score))+
  stat_boxplot(geom = "errorbar", width = 0.15)+
  geom_boxplot()+
  scale_x_discrete("Movie has ever in the Top 200 Box Office", labels = c("No", "Yes"))+
  scale_y_continuous("Audience Score")
g26
```

```{r}
# Distrbution of Audience Score by Audience rating of the Movie
g27 <- ggplot(movies2, aes(x = audience_rating, y = audience_score))+
  stat_boxplot(geom = "errorbar", width = 0.15)+
  geom_boxplot()+
  scale_x_discrete("Audience rating of the Movie")+
  scale_y_continuous("Audience Score")
g27
```

```{r}
# Distribution of Audience Score by Type of Movie
g28 <- ggplot(movies2, aes(x = title_type, y = audience_score))+
  stat_boxplot(geom = "errorbar", width = 0.15)+
  geom_boxplot()+
  scale_x_discrete("Type of the Movie")+
  scale_y_continuous("Audience Score")
g28
```

```{r}
# Distribution of Response Variable by selected Categorical Variables
grid.arrange(g19, g20)
```

```{r}
grid.arrange(g21, g22, g26, g27, ncol = 2)
```

```{r}
grid.arrange(g23, g24, g25, g28, ncol = 2)
```

```{r}
# Distribution of Audience Score by Critics rating
g29 <- ggplot(movies2, aes(x = critics_rating, y = audience_score))+
  stat_boxplot(geom = "errorbar", width = 0.15)+
  geom_boxplot()+
  scale_x_discrete("Critics Rating of the Movie")+
  scale_y_continuous("Audience Score")
g29
```

```{r}
# Audience Score grouped by Movies ever received an Oscar Best Picture Nomination
# Summary Statistics Table
by(movies2$audience_score, movies2$best_pic_nom, summary)
```

```{r}
# Test equal variance
var.test(movies2$audience_score~movies2$best_pic_nom, alternative = "two.sided") # unequal variance
```

```{r}
# Two-Sample T-Test
t.test(audience_score~best_pic_nom, data = movies2)
```

```{r}
# Audience score grouped by Movies ever won an Oscar Best Picture: Summary Statistics Table
by(movies2$audience_score, movies2$best_pic_win, summary)
```

```{r}
# Equal Variance Test
var.test(audience_score~best_pic_win, data = movies2, alternative = "two.sided")
```

```{r}
# Two-sample T-test
t.test(audience_score~best_pic_win, data = movies2)
```

```{r}
# Audience score grouped by Movies ever in the Top 200 Box Office: Summary Statistics Table
by(movies2$audience_score, movies2$top200_box, summary)
```

```{r}
# Equal Variance Test
var.test(audience_score~top200_box, data = movies2, alternative = "two.sided") # cannot reject H0
```

```{r}
# Two-sample T-test
t.test(audience_score~top200_box, data = movies2, var.equl = TRUE)
```

```{r}
# Audience score grouped by Audience rating on Rotten Tomatoes: Summary Statistics Table
by(movies2$audience_score, movies2$audience_rating, summary)
```

```{r}
# Equal Variance Test
var.test(audience_score~audience_rating, data = movies2, alternative = "two.sided")
```

```{r}
# Two-Sample T-test
t.test(audience_score~audience_rating, data = movies2)
```

```{r}
# Audience score grouped by Actors ever won an Oscar: Summary Statistics Table
by(movies2$audience_score, movies2$best_actor_win, summary)
```

```{r}
# Equal Variance Test
var.test(audience_score ~ best_actor_win, data = movies2, alternative = "two.sided") # cannot reject H0
```

```{r}
# Two-Sample T-test
t.test(audience_score~best_actor_win, data = movies2, var.equl = TRUE)
```

```{r}
# Audience Score grouped by Actresses ever won an Oscar: Summary Statistics Table
by(movies2$audience_score, movies2$best_actress_win, summary)
```

```{r}
# Equal Variance Test
var.test(audience_score ~ best_actress_win, data = movies2, alternative = "two.sided") 
# cannot reject H0
```

```{r}
# Two-Sample T-Test
t.test(audience_score~best_actress_win, data = movies2, var.equl = TRUE)
```

```{r}
# Audience Score grouped by Movie Director has ever won an Oscar: Summary Statistics Table
by(movies2$audience_score, movies2$best_dir_win, summary)
```

```{r}
# Equal Variance Test
var.test(audience_score ~ best_dir_win, data = movies2, alternative = "two.sided") 
# cannot reject H0
```

```{r}
# Two-Sample T-Test
t.test(audience_score~best_dir_win, data = movies2, var.equal = TRUE)
```

```{r}
# Audience Score grouped by Movie Type: Summary Statistics Table
by(movies2$audience_score, movies2$title_type, summary)
```

```{r}
# Audience Score grouped by Critics Rating: Summary Statistics Table
by(movies2$audience_score, movies2$critics_rating, summary)
```

```{r}
as_certified_fresh <- movies2 %>%
  filter(critics_rating == "Certified Fresh") %>%
  dplyr::select(audience_score)
as_rotten <- movies2 %>%
  filter(critics_rating == "Rotten") %>%
  dplyr::select(audience_score)
```

```{r}
# Equal Variance Test
var.test(as_certified_fresh$audience_score, as_rotten$audience_score, alternative = "two.sided")
```

```{r}
# Two Sample T-Test
t.test(as_certified_fresh$audience_score, as_rotten$audience_score)
```

# 2. Modelling

# 2.1 Multicollinearity

```{r}
# Correlation Plot between numerical variables
ggpairs(movies2, columns = c("runtime", "imdb_rating", "imdb_num_votes", "critics_score"),
        progress = FALSE)
```

```{r}
correlation.matrix <- cor(movies2[, c("runtime", "imdb_rating", "imdb_num_votes", "critics_score")])
correlation.matrix
```

```{r}
stargazer(correlation.matrix, title = "Correlation Table for Numerical Variables")
```

There exists strong positive relationship between movie's IMDB rating and critics score on Rotten Tomatoes. The correlation 
between those predictors is 0.765. This could result in multicollinearity issue. Therefore, removing one of those two variables 
could improve our estimates. To decide which of those two predictors will be removed from the analysis, I decide to take a look 
at the relationship between audience score (response variable) and IMDB rating, as well as critics score.

```{r}
# Relationship between Audience Score and IMDB Rating
g30 <- ggplot(movies2, aes(x = imdb_rating, y = audience_score))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)+
  scale_x_continuous("IMDB Rating")+
  scale_y_continuous("Audience Score")+
  ggtitle("Scatterplot of IMDB Rating vs. Audience Score")+
  theme(plot.title = element_text(size = 12, face = "bold", hjust = 0.5))
g30
```

```{r}
reg1 <- lm(audience_score ~ imdb_rating, data = movies2)
summary(reg1)
```

```{r}
# Relationship between Audience Score and Critics Score
g31 <- ggplot(movies2, aes(x = critics_score, y = audience_score))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)+
  scale_x_continuous("Critics Score")+
  scale_y_continuous("Audience Score")+
  ggtitle("Scatterplot of Critics Score and Audience Score")+
  theme(plot.title = element_text(size = 12, face = "bold", hjust = 0.5))
g31
```

```{r}
reg2 <- lm(audience_score ~ critics_score, data = movies2)
summary(reg2)
```

The $R^2$ for the estimated regression model between audience score and IMDB rating is 0.7481, which is higher compared to the 
$R^2$ for the regression model between audience score and critics score. Therefore, I decide to keep IMDB rating in the 
analysis.

```{r}
# Create a data set for modeling
movies3 <- movies2 %>%
  dplyr::select(title_type, genre, runtime, mpaa_rating, imdb_rating, imdb_num_votes,
                critics_rating, audience_rating, audience_score, best_pic_nom, best_pic_win,
                best_actor_win, best_actress_win, best_dir_win, top200_box)
head(movies3, 5)
```

# 2.2 Forward Stepwise Selection with AIC

```{r}
# Forward Stepwise with AIC
min.model <- lm(audience_score ~ 1, data = movies3) # start with only intercept
max.model <- lm(audience_score ~ ., data = movies3) # stop with model with everything
scp <- list(lower = min.model, upper = max.model)
fwd <- stepAIC(min.model, direction = "forward", scope = scp)
```

The forward stepwise slection with AIC identified 6 covariates, which are IMDB rating, audience rating on Rotten Tomatoes, 
genre, movie ever received an Oscar Best Picture nomination, movie's runtime, and movie's actress ever won an Oscar award.

```{r}
reg3 <- lm(audience_score ~ imdb_rating + audience_rating + genre + best_pic_nom + 
             runtime + best_actress_win,
           data = movies3)
summary(reg3)
```

```{r}
# Residuals vs. Fitted Values
g32 <- ggplot(data = reg3, aes(x = .fitted, y = .resid))+
  geom_point()+
  geom_hline(yintercept = 0, linetype = "dashed")+
  scale_x_continuous("Fitted Values")+
  scale_y_continuous("Residuals")
g32
```

```{r}
bptest(reg3)
```

# 2.3 Forward Stepwise Selection with BIC

```{r}
# Forward Selection with BIC (k=log(N)), StepAIC function can do the same job for BIC
fwd2 <- stepAIC(min.model,
                direction = "forward",
                scope = scp,
                k = log(nrow(movies3)))
```

Using the Forward Stepwise Selection with BIC identified 2 covariates, which are IMDB rating and 
audience rating on Rotten Tomatoes.

```{r}
reg4 <- lm(audience_score ~ imdb_rating + audience_rating, data = movies3)
summary(reg4)
```

The Adjusted R-squared for the BIC is 0.8821, while the Adjusted R-squared for the AIC is 0.8861. Although the latter is higher,
I would consider using the BIC. This is due to the fact that AIC adds 4 more covariates into the model but only improved 0.004 
(very small) in terms of Adjusted R-Squared. BIC favors more parsimonious model than AIC and it produces a more easily 
interpretable model. While using the model suggested by the AIC does make good predictions, it requires more information. 

# 2.4 Model Diagnostics

Let's do model diagnostics for the model suggested by BIC.

# 2.4.1 Outliers, Leverage Vlues and Influential Points

```{r}
# Get Influence Plot
influencePlot(reg4,  main = "Influence Plot",
              sub = "Circle size is proportional to Cook's distance")
```

Test for Outliers using the Bonferroni outlier test.

```{r}
# Bonferroni p-value for most extreme observations
# The null for this test is that the observation is not an outlier
outlierTest(reg4, cutoff = 0.05)
```

Observation 126 is the most extreme observation in our data, but it isn't an outlier, as the null cannot
be rejected.

\vspace{12pt}

```{r}
# Use Build-in function to get influence measures
im <- influence.measures(reg4)
influential_measures <- as.data.frame(im$infmat)
head(influential_measures, 5)
```

Leverage Values: Large leverage values are greater than $2p/n$.

```{r}
leverage_cutoff <- 2*(3/650)
(which(influential_measures$hat > leverage_cutoff))
```

```{r}
(which(im$is.inf[, 7] == "TRUE"))
```


```{r}
# Prepare a Leverage Plot
leverage_data <- im$infmat[, 7]
leverage_influence <- im$is.inf[, 7]
index <- seq(1, 650, by = 1)
leverage_data <- as.data.frame(cbind(leverage_data, leverage_influence, index))
leverage_data <- leverage_data %>%
  dplyr::select("hat_values" = leverage_data,
                "INFLUENCE" = leverage_influence,
                index)
head(leverage_data, 5)
```

```{r}
gleverage <- ggplot(leverage_data, aes(x = as.factor(index), 
                                       y = hat_values, label = index))+
  geom_point()+
  geom_text(data = subset(leverage_data, INFLUENCE ==1), nudge_x = 0.05, vjust = -0.5, size = 3)+
  scale_x_discrete("Case Index", breaks = seq(0, 650, by = 100))+
  scale_y_continuous("Hat Values")+
  geom_hline(yintercept = leverage_cutoff, linetype = "dashed")
gleverage
```

\vspace{12pt}

DFFITS: cut off rule: greater than $2\sqrt{\frac{p}{n}}$ for data set with observations $n \geqq 40$.

```{r}
# Below is how Build-in function find those points
which(im$is.inf[, 4] == "TRUE")
```

```{r}
# Prepare an DFFITS plot
dffits_data <- abs(im$infmat[, 4])
dffits_influence <- im$is.inf[, 4]
dffits_combined <- as.data.frame(cbind(dffits_data, dffits_influence))
dffits_data <- cbind(dffits_combined, index)
dffits_data <- dffits_data %>%
  dplyr::select("DFFITS" = 'dffits_data',
                "INFLUENCE" = 'dffits_influence',
                index)
```

```{r}
gdffits <- ggplot(dffits_data, aes(x = as.factor(index), y = DFFITS, label = index))+
  geom_point()+
  geom_text(data = subset(dffits_data, INFLUENCE == 1), nudge_x = 0.05, vjust = -0.5, size = 3)+
  scale_x_discrete("Case Index", breaks=seq(0,650, by = 100))+
  scale_y_continuous("DFFITS")+
  geom_hline(yintercept = 2*sqrt(3/650), linetype = "dashed")
gdffits
```

\vspace{12pt}

Cook's Distance: If Cooks distance > 4/n the observation is considered too influential

```{r}
# Identify Influential Points use Cook's Distance
cook_cutoff <- 4/nrow(movies3)
plot(reg4, which = 4, cook.levels = cook_cutoff)
```

```{r}
# Check influential points' Cook's distance: we have 29 observations larger than cutoff point
cooks_distance_df <- influential_measures %>%
  dplyr::select(cook.d) %>%
  filter(cook.d > cook_cutoff) %>%
  arrange(desc(cook.d))
cooks_distance_df
```

If you use another cutoff value: $F(0.50; p, n-p)$

```{r}
cook_cutoff2 <- qf(0.50, 3, 647)
which(im$infmat[, 6] > cook_cutoff2)
```

```{r}
which(im$is.inf[, 6] == "TRUE")
```

This returns that I don't see any cases that are very influential based on Cook's distance.

# 2.4.2 Heteroskedasticity

```{r}
# Linear Relationship between Numerical Variable X and Y: Functional Form is correct
olsresiduals <- residuals(reg4)
ggplot(movies3, aes(x = imdb_rating, y = olsresiduals))+
  geom_point()+
  scale_x_continuous("IMDB Rating")+
  scale_y_continuous("Residuals")+
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
# Residuals vs. Fitted Values
g33 <- ggplot(data = reg4, aes(x = .fitted, y = .resid))+
  geom_point()+
  geom_hline(yintercept = 0, linetype = "dashed")+
  scale_x_continuous("Fitted Values")+
  scale_y_continuous("Residuals")
g33
```

The residuals vs. fitted values plot exhibits a megaphone shape. There is more variation in the lower level of fitted audience scores. Confirm heteroskedasticity. 

```{r}
# Breusch Pagan Test for heteroskedasticity
bptest(reg4)
```

We can also use Breusch-Pagan test. P-value is very small, we should reject the null hypothesis and conclude heteroskedasticity.

```{r}
y_hat <- predict(reg4)
residuals.stdres <- rstandard(reg4)
data <- as.data.frame(cbind(y_hat, residuals.stdres))
```

```{r}
# Standardized Residuals vs. Fitted Values
g34 <- ggplot(data, aes(x = y_hat, y = residuals.stdres))+
  geom_point()+
  geom_hline(yintercept = 0, linetype = "dashed")+
  scale_x_continuous("Fitted Values")+
  scale_y_continuous("Standardized Residuals")
g34
```

# 2.4.3 Normal Probability of Residuals

```{r}
# Normal Probability of Residuals
hist_res <- ggplot(data = reg4, aes(x = .resid))+
  geom_histogram(fill = "white", col = "black", bins = 50)+
  geom_vline(xintercept = median(reg4$residuals), linetype = "dashed", col = "blue")+
  geom_vline(xintercept = mean(reg4$residuals), linetype = "dashed", col = "red")+
  geom_text(label = "Median", x = 4, y = 45, col = "blue")+
  geom_text(label = "Mean", x = -2, y = 45, col = "red")+
  xlab("Residuals")+
  ylab("Count")+
  ggtitle("Histogram of Residuals")+
  theme(plot.title = element_text(size = 10, hjust = 0.5, face = "bold"))
hist_res
```

```{r}
qq_res <- ggplot(data = reg4, aes(sample = .resid))+
  stat_qq()+stat_qq_line()+
  ggtitle("Normal Q-Q Plot")+
  theme(plot.title = element_text(size = 10, face = "bold", hjust = 0.5))
qq_res
```
```{r}
grid.arrange(hist_res, qq_res, ncol = 2)
```

```{r}
# Anderson-Darling Test for Normality
ad.test(reg4$residuals) # The data does not follow a normal distribution
```

```{r}
# Shapiro-Wilk
shapiro.test(reg4$residuals)
```

# 2.4.4 Independent Residuals

```{r}
# Assumption: Independent Residuals
dwtest(reg4) # p-value = 0.2705, Cannot reject H0
```

# 2.4.5 Multicollinearity use VIF

```{r}
# Check Multicollinearity
vif(reg4)
```

The VIF are all smaller than 5, there is no multicollinearity issue after we use OLS model. 

To sum up, we should deal with heteroskedasticity issue..

# 3. Feasible Generalized Least Squares (FGLS)

We do not know the covariance of variance. So, let's use FGLS method

```{r}
# Squared Residuals vs. Fitted Values
residuals2 <- residuals(reg4)^2
g35 <- ggplot(reg4, aes(x = .fitted, y = residuals2))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)+
  scale_x_continuous("Fitted Values")+
  scale_y_continuous("Squared Residuals")
g35
```

```{r}
# Squared Residuals vs. IMDB Rating
g36 <- ggplot(movies3, aes(x = imdb_rating, y = residuals2))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)+
  scale_x_continuous("IMDB Rating")+
  scale_y_continuous("Squared Residuals")
g36
```

```{r}
lne2 <- log(residuals2)
reg5 <- lm(lne2~imdb_rating+audience_rating, data = movies3)
summary(reg5)
```

```{r, results='hide'}
# Obtain estimated variances
lne2fitted <- predict(reg5)
v <- exp(lne2fitted)
```

```{r, results='hide'}
# Estimate weight for each case
w <- 1/v
```

```{r}
# Feasible Generalized Least Squares
reg.gls <- lm(audience_score ~ imdb_rating + audience_rating, weights = w, data = movies3)
summary(reg.gls)
```

```{r}
y_hat2 <- predict(reg.gls)
gls_stdres <- rstandard(reg.gls)
data3 <- as.data.frame(cbind(y_hat2, gls_stdres))
```

```{r}
# FGLS: Standardized Residuals vs. Fitted Values
g37 <- ggplot(data3, aes(x = y_hat2, y = gls_stdres))+
  geom_point()+
  geom_hline(yintercept = 0, linetype = "dashed")+
  scale_x_continuous("Fitted Values")+
  scale_y_continuous("Standardized Residuals")
g37
```

Using Feasible GLS, according to the the Standardized Residuals vs. Fitted Values plot, it looks much better (Although 
heteroskedasticity may still exists, it is much better improved) and it has a higher Adjusted R-Squared (0.889). The FGLS model 
is definitely more precise.

```{r}
# Test for Normality of Residuals
ad.test(reg.gls$residuals)
```

\newpage

# 4. Prediction

I am going to predict the audience score for a new movie that has not been used to fit the model. The movie I choose is "The 
Incredible Hulk". A movie produced by Universal Pictures in 2008. I extracted the audience rating from Rotten Tomatoes and IMDB 
rating from IMDB (6.7, which lies inside the range of the values of the predictor variables in our model). The true audience 
score for this movie is 70.

```{r}
newmovie <- as.data.frame(cbind(imdb_rating = 6.7, 
                         audience_rating = "Upright"))
newmovie <- newmovie %>%
  mutate(imdb_rating = as.numeric(imdb_rating))
```

```{r}
# Use OLS to predict
predict(reg4, newmovie, interval = "prediction")
```

The OLS model predicts movie "The Incredible Hulk" will have an audience score at approximately 73.16. The model predicts, with 95% confidence, that the movie is expected to have an audience score between 59.49 and 86.83. The prediction interval contains the actual audience score for this movie.

\vspace{12pt}

```{r}
# Obtain OLS residual
y_hulk <- 70
(olse <- y_hulk - predict(reg4, newmovie))
(olse2 = olse^2)
```

```{r}
# Use Feasible GLS to predict
# First, we have to calculate weight for this case
data4 <- as.data.frame(cbind(lne2 = log(olse2),
                             imdb_rating = 6.7,
                             audience_rating = "Upright"))
data4 <- data4 %>%
  mutate(imdb_rating = as.numeric(imdb_rating))
```

```{r}
lne2fitted2 <- predict(reg5, data4)
v2 <- exp(lne2fitted2)
(w2 <- 1/v2)
```

```{r}
predict(reg.gls, newmovie, interval = "prediction", weights = w2)
```

Again, the FGLS confidence interval is narrower than OLS. This is another way to show that FGLS is more efficient than OLS.

# 5. Report Regression Results

```{r, results='hide'}
# Generate Regression Tables in LaTex form
stargazer(reg4, reg.gls, title = "Regression Results",
          dep.var.labels = c("Audience Score"),
          covariate.labels = c("IMDB Rating", "Audience Rating"),
          column.labels = c("OLS", "FGLS"))
```

\newpage

# 6. Optional: Iterative Feasible GLS (IFGLS)

```{r}
# We can use a function to do the IFGLS
test <- vector('list', 50) # Set limits as 50
test[[1]] <- list(coef0 = coef(reg4), res0 = reg4$residuals, 
                  fitted0 = reg4$fitted.values, summary = summary(reg4)) # Begin with OLS
```

```{r}
ifgls_function <- function(obj){
  residuals2 <- obj[['res0']]^2
  logresiduals2 <- log(residuals2)
  residuals2_lm <- lm(logresiduals2~imdb_rating+audience_rating, data = movies3)
  residuals2_fitted <- predict(residuals2_lm)
  estiamted_variance <- exp(residuals2_fitted)
  w <- 1/estiamted_variance
  reg_fgls <- lm(audience_score~imdb_rating+audience_rating, weights = w, data = movies3)
  return(list(coef0 = coef(reg_fgls), res0 = reg_fgls$residuals, 
              fitted0 = reg_fgls$fitted.values, summary = summary(reg_fgls)))
}
```

```{r}
# After iterating 15 times, the coefficients become stable
for (i in 1: 20){
  test[[i+1]] <- ifgls_function(test[[i]])
  cat(i, test[[i]][['coef0']], '\n')
}
```

\vspace{12pt}

Using the Iterative FGLS, the final coefficient of IMDB rating becomes stable at 12.41811 and the coefficient for Audience Rating becomes stable at 17.36031.

```{r}
test[[15]][['summary']]
```


