---
title: "5126HW6"
author: "Chengxuan Hu"
date: "10/14/2020"
output: word_document
---

7.7 on p. 289
```{r}
 # Read Table6.18
 t6.18<-read.table("http://users.stat.ufl.edu/~rrandles/sta4210/Rclassnotes/data/textdatasets/KutnerData/Chapter%20%206%20Data%20Sets/CH06PR18.txt", header=FALSE)
 # Column names can be changed with
 names(t6.18)[1]<-paste("Y")
 names(t6.18)[2]<-paste("X1")
 names(t6.18)[3]<-paste("X2")
 names(t6.18)[4]<-paste("X3")
 names(t6.18)[5]<-paste("X4")
 
 # To use the column names without reference to t6.18 you need to attach the dataset:
 attach(t6.18)
```

a Obtain the analysis of variance table that decomposes the regression sum of squares into extra sums of squares associated with X4; with X1 given X4; with X2, given X1, and X4; and with X3, given X1 X2 and X4.
```{r}
reg4123 = lm(Y ~ X4 + X1 + X2 + X3)
anova(reg4123)
# SSR(X4) = 67.775
# SSR(X1|X4) = 42.275
# SSR(X2|X1,X4) = 27.857
# SSR(X3|X1,X2,X4) = 0.420
```

b. Test whether X3 can be dropped from the regression model given that X1, X2 and X4 are retained.
Use the F* test statistic and level of significance .01. State the alternatives, decision rule, and conclusion. What is the p-value of the test?
```{r}
reg412 = lm(Y ~ X4 + X1 + X2)
anova(reg412, reg4123)

qf(0.99,1,76)

# H0: b3 = 0 
# Ha: b3 ≠ 0
# Decision rule: if F*>critical value, reject H0 
              ## if p-value < 0.01, reject H0
# Conlusion: F*=0.3248 < qf(0.99,1,76)=6.980578, do not reject H0, that is X3 can be dropped from the regression model given that X1, X2 and X4 is retained.
# p-value = 0.5704 > 0.01, do not reject H0
```

Refer to Commercial properties Problems 6.18 and 7.7. Calculate R^2(Y4), R^2(Y1),  R^2(Y1|4), R^2(14), R^2(Y2|14), R^2(3|124), and R^2. Explain what each coefficient measures and interpret your results. How is the degree of marginal linear association between Y and X1 affected, when adjusted for X4?
```{r}
library(rsq)

reg4 = lm(Y ~ X4)
rsq.partial(reg4)
# R^2(Y4)=0.2865058, that is proportion of Variation in Y explained by X4 is 28.65%

reg1 = lm(Y ~ X1)
rsq.partial(reg1)
# R^2(Y1)=0.06264236, that is proportion of Variation in Y explained by X1 is 6.26%

reg41 = lm(Y ~ X4 + X1)
rsq.partial(reg41, reg4)
# R^2(Y1|4)=0.2504679, that is proportion of Variation in Y not explained by X4 , that is explained by X1 is 25.05%

(cor(X1,X4))^2
# R^2(14)=0.08328044, that is the square of the correlation between X1 and X4 is 0.08. They are weakly correlated.

reg412 = lm(Y ~ X4 + X1 + X2)
rsq.partial(reg412, reg41)
# R^2(Y2|14)=0.2202037, that is proportion of Variation in Y not explained by X4 and X1 but explained by X2 is 22.02%

rsq.partial(reg4123, reg412)
# R^2(Y3|124)=0.004254889, that is proportion of Variation in Y not explained by X4, X1 and X2 but explained by X3 is 0.43%

summary(reg4123)
# R^2=0.5847, that is proportion of Variation in Y explained by X1, X2, X3and X4 is 58.47%
```

How is the degree of marginal linear association between Y and X1 affected, when adjusted for X4?
```{r}
# R^2(Y1)=0.06264236
# R^2(Y1|4)=0.2504679
# Therefore, the degree of marginal linear association between Y and X1 increases when adjusted for X4.
```

